Verification of a Deterministic Random Bits Generator

### 2015-12-27 paper/paper.tex
added explicitly that HMAC is irreversible
more tweaks
added a section on proof process

\usepackage{hyperref}
\usepackage{couriers}
\usepackage[normalem]{ulem}
\usepackage{listings}
\usepackage{lstlangcoq}
\usepackage{appendix}

\lstset{language=Coq,basicstyle=\ttfamily,mathescape=true,columns=fullflexible,breaklines=true,breakatwhitespace=true}
\hypersetup{citecolor=black,linkcolor=black}

@@

\stdtitle{NIST SP 800-90A} specifies different types of DRBG. The version I verified in this project is the one based on the key-hashed message authentication code, or HMAC. HMAC takes a (secret) key and a message and computes a fixed-length message authentication code. In theory, someone with only the message will not be able to compute the message authentication code, and someone with both the message and the message authentication code will not be able to compute the key.

@@

\subsection{Modifications to mbedTLS}
I had to make a few modifications to the \stdtitile{mbedTLS} DRBG source file. Most of these transformations are obviously equivalent to the original version. For example, Verifiable C does not allow memory dereferences in function parameters or if/while/for-loop conditions.

Therefore, code such as the following had to be changed:

@@

to this:

@@

\subsection{The Update Function}
Before showing any of the public function, I must first define a helper function, the \lstinline{update} function. See \stdtitle{NIST SP 800-90A \S 10.1.2.2} for the corresponding NIST specification. The \lstinline{update} function takes a \lstinline{key}, a \lstinline{value}, and an optional nonce and returns a new \lstinline{key} and a new \lstinline{value}. This function will be called at least once in each of the public functions, and the goal is to provide backtracking resistance. This is achieved by having the secret internal state of the DRBG (the \lstinline{key}) be the output of HMAC.

@@

\subsection{A Description of an API Specification}

@@

Both the pre- and post-conditions are of the form \lstinline{PROP () LOCAL () SEP ()}. This is separation logic. As an example, take the following:

@@

@@

I verified the \stdtitle{mbedTLS} implementations of \lstinline{update}, \lstinline{reseed}, and \lstinline{generate} correct with respect to their API specifications. The actual proof is long and tedious. Fortunately, the proof is machine-checked in Coq, and so can be trusted to be correct. I will not present the proof here. Instead, I will provide some statistics.

>>>>

\usepackage{hyperref}
\usepackage{couriers}
\usepackage[normalem]{ulem}
\usepackage{listings}
\usepackage{lstlangcoq}
\usepackage{appendix}
\usepackage{graphicx}

\lstset{language=Coq,basicstyle=\ttfamily,mathescape=true,columns=fullflexible,breaklines=true,breakatwhitespace=true}
\hypersetup{citecolor=black,linkcolor=black}
\graphicspath{ {images/} }

@@

\stdtitle{NIST SP 800-90A} specifies different types of DRBG. The version I verified in this project is the one based on the key-hashed message authentication code, or HMAC. HMAC takes a (secret) key and a message and computes a fixed-length message authentication code. In theory, someone with only the message will not be able to compute the message authentication code, and someone with both the message and the message authentication code will not be able to compute the key. Note that HMAC is an irreversible function.

@@

\subsection{Modifications to mbedTLS}
I had to make a few modifications to the \stdtitle{mbedTLS} DRBG source file. Most of these transformations are obviously equivalent to the original version. For example, Verifiable C does not allow memory dereferences in function parameters or if/while/for-loop conditions. Therefore, code such as the following had to be changed:

@@

\noindent to this:

@@

\subsection{The Update Function}
Before showing any of the public function, I must first define a helper function, the \lstinline{update} function. See \stdtitle{NIST SP 800-90A \S 10.1.2.2} for the corresponding NIST specification. The \lstinline{update} function takes a \lstinline{key}, a \lstinline{value}, and an optional nonce and returns a new \lstinline{key} and a new \lstinline{value}. This function will be called at least once in each of the public functions, and the goal is to provide backtracking resistance. This is achieved by having the secret internal state of the DRBG (the \lstinline{key}) be the output of HMAC, an irreversible function.

@@

\subsection{A Description of an API Specification}\label{api_spec}

@@

Both the pre- and post-conditions are of the form \lstinline{PROP() LOCAL() SEP()}. This is separation logic. As an example, take the following:

@@

\subsection{Proof Process}

In this section, I will explain what a typical proof process looks like. I will not be showing complete proofs that the \stdtitle{mbedTLS} functions conform to the API specifications, as such proofs are long and tedious. Fortunately, the proofs are machine-checked in Coq, and so can be trusted to be correct.
\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{before_forward}
\caption{The \lstinline{update} Function Proof}
\label{figure:before_forward}
\end{figure}

Figure \ref{figure:before_forward} shows a part of the \lstinline{update} function proof. The left pane shows the actual proof, while the right pane shows the current proof goal.

On the left pane, everything highlighted in blue is the part of the proof that has been accepted by the Coq compiler as correct. Everything in white has not been accepted, and so can still be modified. Completing the proof means compiling the entire file, or turning the entire file blue.

In the current proof goal is the proof environment, of the same \lstinline{PROP() LOCAL() SEP()} form as seen previously in \S \ref{api_spec}. The proof environment contains propositional facts, facts about local variables, and facts about memory.

The current proof goal also contains the abstract syntax tree (AST) of the C code. Represented as nodes in the AST, each statement modifies the proof environment, and each step in the proof (left pane) is an instruction to Coq on how to step through the AST. For simple statements, such as an assignment statement, the proof automation system can figure out how the proof environment will be affected by stepping through the statement. Other statements require the proof to explicitly state how the proof environment will be affected. These statements include the \lstinline{if} statement or different kinds of loops.

Figure \ref{figure:before_forward} shows the case where the next statement in the AST is an assignment statement. This is the simple case, and so the proof command \lstinline{forward} can automatically figure out how the environment will be modified. Figure \ref{figure:after_forward} shows the same proof after the \lstinline{forward} command has been accepted by the Coq compiler. Notice the proof environment has changed in the \lstinline{LOCAL} part, and the AST no longer shows the assignment statement as the next statement.

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{after_forward}
\caption{After Compiling the \lstinline{forward} Command}
\label{figure:after_forward}
\end{figure}

The proof environment is initialized to the function preconditions. A complete proof shows that, after the entire AST has been stepped through, the modified environment implies the function postconditions. Figure \ref{figure:qed} shows when the entire proof has been compiled. Stepping through the \lstinline{Qed} command will complete the proof.

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{qed}
\caption{At the End of the \lstinline{update} Proof}
\label{figure:qed}
\end{figure}

@@

I verified the \stdtitle{mbedTLS} implementations of \lstinline{update}, \lstinline{reseed}, and \lstinline{generate} correct with respect to their API specifications. In this section, I will provide some statistics about the verifications.
